Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
local changes

\noindent
In the big-data world, similar information is often stored or transmitted in
different formats. For instance, electronic calendars come in iCalendar,
vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can
be transmitted via Tax XML (used by the US government for e-filing) or TXF
(used by TurboTax)~\cite{tax-formats}; and sewing machines can use
embroidery files in formats with names like PES, DST, and several
others~\cite{embroidery-formats}.  \bcp{It would be nice to be able to give
  a quick example of a truly symmetric case---do any of these examples work
  for that?}

@@

The first difficulty is pragmatic. Symmetric lenses usually operate
over three structures: the source $S$, the target $T$ and a complement
$C$ that contains the information not
present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
somewhere and managed. More significantly, complements complicate synthesis
specifications.
Users can no longer just give examples of source and target
pairings; instead they must also indicate how to manage the 
``hidden state.'' To avoid such complexities, we define a new variant
of symmetric lenses: \emph{simple symmetric lenses}. At the price
of some expressiveness, these new lenses do not require complements.
%only the usual source and target.
%
To characterize their expressive power,
we prove that simple symmetric lenses include all (and only)
those symmetric lenses that satisfy an intuitive property called
\emph{forgetfulness}.
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $S$ or $T$ when making a round trip. They only need the most
recent instance. 
We justify this loss of expressiveness pragmatically. 
We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
successfully use them on a range of real-world applications, showing
in the process that simple symmetric lenses integrate smoothly with 
other advanced lenses in Boomerang, including quotient
lenses and matching lenses.

@@

With simple symmetric lenses and our SRE-based likeliness measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likeliness measure.
However, regular expressions enjoy a rich algebraic structure, meaning
there are infinitely many regular expressions equivalent to the
original pair, and the lens returned by a type-directed search depends
upon the chosen representation. 
To tame this complexity, we divide the synthesis algorithm into two communicating
search procedures (Figure~\ref{fig:high-level-algorithm}), following
earlier work~\cite{maina+:quotient-synthesis}\bcp{Is this the right citation?}.
The first algorithm, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second algorithm, \GreedySynth, takes a pair of such SREs as input
and uses a greedy, type-directed algoithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The algorithm terminates when the likilihood score of the
best lens found so far compared with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

@@

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}
\begin{itemize}
\item We define {\em simple symmetric lenses}, with the
  the useful property that they do not maintain any hidden state
  (\S\ref{sec:overview}).  
  We present a collection of combinators for defining these lenses and
  describe how to 
  calculate their likelihoods (\S\ref{sec:ssl}). 
\item  We show that simple symmetric lenses
  fall between asymmetric lenses~\cite{Focal2005-long2} and full
  symmetric lenses~\cite{symmetric-lenses} in expressiveness
  (\S\ref{sec:related}).  We further show that the class of simple
  lenses can be characterized semantically as the subset of full
  symmetric lenses for which a simple ``forgetfulness'' property holds.
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
  To support that effort, we show that the star semi-ring equivalences on
  regular 
  expressions remain valid when applied to stochastic regular
  expressions, preserving the probability distributions as well as the languages
  (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with 
  other lens extensions such as quotient lenses~\bcp{citation} and matching
  lenses~\bcp{citation}.  We evaluate our
  implementation on a set of 48 lens synthesis benchmarks drawn from
  data cleaning, view maintenance, 
  and file synchronization tasks.  We demonstrate that we can
  synthesize simple symmetric lenses for all of the benchmarks in
  under 30 seconds (\S\ref{sec:evaluation}). 
\end{itemize}
We close with a discussion of related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

Now that we have a way to calculate likelihood, we need a way to search for
likely lenses. Previous work on synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16} has shown that we can use
types to dramatically restrict the search space and improve the effectiveness of
synthesis. In our setting, types are pairs of stochastic regular expressions,
which come with a rich algebraic structure that makes each regular expression
semantically equal to infinitely many others. To tame that search space and
leverage the benefits of type-directed synthesis, we split our algorithm into
two communicating synthesizers following earlier
work~\cite{maina+:quotient-synthesis}\bcp{Some of this text seems to be
  identical to text in the intro.  That will annoy careful readers.}. The
first synthesizer, \Expand, navigates 
the space of semantically equivalent regular expressions by applying rewrite
rules that preserve both the semantics and the probability distributions of
those stochastic regular expressions. Our algorithm ranks pairs of stochastic
regular expressions by the number of rewrite rule applications required to
obtain the pair from the stochastic regular expressions given as input. It
passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

@@

\begin{figure}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
  \caption{Stochastic Regular Expression Star-Semiring Equivalence\bcp{Can
      we make the skips between the separate sections of the definition a
      bit larger, and increase the margin on the left?  It looks (even)
      denser than it is, at the moment.}}
  \label{fig:ssr-sre}
\end{figure}

@@

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions. We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations. In a slight abuse of notation\saz{Why is this an
  abuse?}\afm{We also use \BRegex{},\BRegexAlt{} as nonterminals}, if $\Regex$
is a stochastic regular expression, $\BRegex$ is the regular expression obtained
by removing probability annotations. The typing judgment $\Lens \OfType \BRegex
\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.

@@

\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
hash-consing stochastic DNF regular expressions.  This optimization enables
fast equality checks and saves memory.\dpw{we could probably cut this paragraph.}
%memoizing its results.  However, our stochastic DNF regular expressions are
%quite large and so hashing them is computationally intensive.  We use less
%memory and allow for quick comparisons and hashing of these DNF regular
%expressions by

@@

For these situations, users can provide a number, $t$, that affects
the strictness of the termination condition, adjusting the \Continue formula
as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
\end{gather*}
With a positive $t$, the termination condition is stricter, and the algorithm
will spend longer searching for more likely lenses. With a negative $t$, the
termination condition is less strict, and the algorithm will spend less time
searching for more likely lenses.  Our default $t$ is $0$.

As discussed in the following section,
while our default termination condition quickly finds the correct lens for
39 of our 48 benchmarks, our termination condition is not
one-size-fits all. Complex lenses, or lenses that project a
large amount of data, can be synthesized more easily and quickly by
adjusting $t$. \dpw{should this paragraph be cut and its content left for the
next section?}
% end synthesis

@@

%\paragraph*{Issues with Classical Symmetric Lenses}
\iffalse Using symmetric lenses for synchronization forces us to maintain not
just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
  unintuitive, and I think we should be careful not to oversell the difficulty
  of storing the complements. It requires more engineering, but it's not rocket
  science.}This is unintuitive, and creates many problems in practice. For
example, consider the situation where I am synchronizing a crontab and job
manager file. I am synchronizing across computers, so I must either maintain a
copy of the complement on each computer, and keep them synchronized, or figure
out a computer to hold the complement, and make sure all edits go through that
complement. Furthermore, if I want to change the synchronization to a different
computer, I don't need to merely make sure that my files are moved to the
computer, I also have to make sure that I correctly move the complement, and
hook up that complement to be used for that synchronization task. This is a huge
divergence from other synchronization utilities\bcp{No. For example, Unison
  maintains a database on each replica recording, for each file, what its hash
  was after the last time the two replicas were synchronized.}, where all that
needs to be done is maintain the files. Using those for synchronization, edits
can be sent over the wire, and the appropriate function can stitch together
those edits, no additional state needed. \fi While classical symmetric lenses
are technically more expressive than simple symmetric lenses, it is unclear
whether this additional expressiveness is useful in practice. In our judgement,
the complications they introduce into synthesis did not justify the potential
additional expressiveness: because each lens has a custom
complement, one can no longer specify the put functions through input/output
examples alone. One alternative would be to enrich specifications with edit
sequentials; another would be to specify the structure of complements explicitly
(though the latter would be somewhat akin to specifying the internal state of a
program). In either case, the complexity of the specifications increases.
%% addition, users would have to specify some auxilary 
%% to specify the behavior through the observable behavior of edit sequences.
%% Second, there are lenses that require arbitrarily long edit sequences to
%% differentiate between: consider the lens that keeps two formats equal for the
%% first $k$ non-identity edits, then keeps them completely unsynchronized for all
%% future edits.  This requires an edit sequence of $k$ edits to differentiate
%% between this lens and the identity lens.  Simple symmetric lenses avoid this
%% problem by disallowing such lenses: we can't keep track of the number of
%% non-identity edits performed unless there are explicit fields in the data
%% formats keeping track of them.  While we are reducing the expressivity of our
%% lenses with this restriction, we feel the benefits outweigh the costs.

>>>>

\noindent
In the big-data world, similar information is often stored or transmitted in
different formats. For instance, electronic calendars come in iCalendar,
vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can be
transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files in
formats with names like PES, DST, and several others~\cite{embroidery-formats}.
\bcp{It would be nice to be able to give a quick example of a truly symmetric
  case---do any of these examples work for that?}\afm{h-event and vCalendar are
  truly symmetric. Even the fields related to location alone are truly
  symmetric (though it is subtle to see this).}

@@

The first difficulty is pragmatic. Symmetric lenses usually operate
over three structures: the source $S$, the target $T$ and a complement
$C$ that contains the information not
present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
somewhere and managed. More significantly, complements complicate synthesis
specifications.
Users can no longer just give examples of source and target
pairings; instead they must also indicate how to manage the 
``hidden state.'' To avoid such complexities, we define a new variant
of symmetric lenses: \emph{simple symmetric lenses}. At the price
of some expressiveness, these new lenses do not require complements.
%only the usual source and target.
%
To characterize their expressive power,
we prove that simple symmetric lenses include all (and only)
those symmetric lenses that satisfy an intuitive property called
\emph{forgetfulness}.
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $S$ or $T$ when making a round trip. They only need the most
recent instance. 
We justify this loss of expressiveness pragmatically. 
We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
successfully use them on a range of real-world applications, showing
in the process that simple symmetric lenses integrate smoothly with 
other advanced lenses in Boomerang, including quotient
lenses~\cite{quotientlenses} and matching lenses~\cite{matchinglenses}.

@@

With simple symmetric lenses and our SRE-based likeliness measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likeliness measure.
However, regular expressions enjoy a rich algebraic structure, meaning
there are infinitely many regular expressions equivalent to the
original pair, and the lens returned by a type-directed search depends
upon the chosen representation. 
To tame this complexity, we divide the synthesis algorithm into two communicating
search procedures (Figure~\ref{fig:high-level-algorithm}), following
earlier work~\cite{optician}.
The first algorithm, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second algorithm, \GreedySynth, takes a pair of such SREs as input
and uses a greedy, type-directed algoithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The algorithm terminates when the likilihood score of the
best lens found so far compared with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

@@

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
  when the contributions mimic the paper, I am not jarred when they don't.}
\begin{itemize}
\item We define {\em simple symmetric lenses}, with the
  the useful property that they do not maintain any hidden state
  (\S\ref{sec:overview}).  
  We present a collection of combinators for defining these lenses and
  describe how to 
  calculate their likelihoods (\S\ref{sec:ssl}). 
\item  We show that simple symmetric lenses
  fall between asymmetric lenses~\cite{Focal2005-long2} and full
  symmetric lenses~\cite{symmetric-lenses} in expressiveness
  (\S\ref{sec:related}).  We further show that the class of simple
  lenses can be characterized semantically as the subset of full
  symmetric lenses for which a simple ``forgetfulness'' property holds.
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
  To support that effort, we show that the star semi-ring equivalences on
  regular 
  expressions remain valid when applied to stochastic regular
  expressions, preserving the probability distributions as well as the languages
  (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with 
  other lens extensions such as quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}.  We evaluate our
  implementation on a set of 48 lens synthesis benchmarks drawn from
  data cleaning, view maintenance, 
  and file synchronization tasks.  We demonstrate that we can
  synthesize simple symmetric lenses for all of the benchmarks in
  under 30 seconds (\S\ref{sec:evaluation}). 
\end{itemize}
We close with a discussion of related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

Now that we have a way to calculate likelihood, we need a way to search for
likely lenses. Previous work on
synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16}
has shown that we can use types to dramatically restrict the search space and
improve the effectiveness of synthesis. In our setting, types are pairs of
stochastic regular expressions where each regular expression is semantically
equal to infinitely many others. We follow existing work on bidirectional
program synthesis, and split our algorithm into two communicating
synthesizers~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand,
navigates the space of semantically equivalent regular expressions by applying
rewrite rules that preserve both the semantics and the probability distributions
of those stochastic regular expressions. Our algorithm ranks pairs of stochastic
regular expressions by the number of rewrite rule applications required to
obtain the pair from the stochastic regular expressions given as input. It
passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

@@

\begin{figure}
  \setlength\extrarowheight{3pt}
  \begin{tabular}{@{\hspace*{5em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
  \label{fig:ssr-sre}
\end{figure}

@@

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions. We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations. In a slight abuse of notation, if $\Regex$ is a
stochastic regular expression, $\BRegex$ is the regular expression obtained by
removing probability annotations. The typing judgment $\Lens \OfType \BRegex
\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.

@@

\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
hash-consing stochastic DNF regular expressions.  This optimization enables
fast equality checks and saves memory.
%memoizing its results.  However, our stochastic DNF regular expressions are
%quite large and so hashing them is computationally intensive.  We use less
%memory and allow for quick comparisons and hashing of these DNF regular
%expressions by

@@

For these situations, users can provide a number, $t$, that affects
the strictness of the termination condition, adjusting the \Continue formula
as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
\end{gather*}
With a positive $t$, the termination condition is stricter, and the algorithm
will spend longer searching for more likely lenses. With a negative $t$, the
termination condition is less strict, and the algorithm will spend less time
searching for more likely lenses.  Our default $t$ is $0$.
% end synthesis

@@

%\paragraph*{Issues with Classical Symmetric Lenses}
%\iffalse Using symmetric lenses for synchronization forces us to maintain not
%just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
%  unintuitive, and I think we should be careful not to oversell the difficulty
%  of storing the complements. It requires more engineering, but it's not rocket
%  science.}This is unintuitive, and creates many problems in practice. For
%example, consider the situation where I am synchronizing a crontab and job
%manager file. I am synchronizing across computers, so I must either maintain a
%copy of the complement on each computer, and keep them synchronized, or figure
%out a computer to hold the complement, and make sure all edits go through that
%complement. Furthermore, if I want to change the synchronization to a different
%computer, I don't need to merely make sure that my files are moved to the
%computer, I also have to make sure that I correctly move the complement, and
%hook up that complement to be used for that synchronization task. This is a huge
%divergence from other synchronization utilities\bcp{No. For example, Unison
%  maintains a database on each replica recording, for each file, what its hash
%  was after the last time the two replicas were synchronized.}, where all that
%needs to be done is maintain the files. Using those for synchronization, edits
%can be sent over the wire, and the appropriate function can stitch together
%those edits, no additional state needed. \fi

While classical symmetric lenses
are technically more expressive than simple symmetric lenses, it is unclear
whether this additional expressiveness is useful in practice. In our judgement,
the complications they introduce into synthesis did not justify the potential
additional expressiveness: because each lens has a custom
complement, one can no longer specify the put functions through input/output
examples alone. One alternative would be to enrich specifications with edit
sequentials; another would be to specify the structure of complements explicitly
(though the latter would be somewhat akin to specifying the internal state of a
program). In either case, the complexity of the specifications increases.
%% addition, users would have to specify some auxilary 
%% to specify the behavior through the observable behavior of edit sequences.
%% Second, there are lenses that require arbitrarily long edit sequences to
%% differentiate between: consider the lens that keeps two formats equal for the
%% first $k$ non-identity edits, then keeps them completely unsynchronized for all
%% future edits.  This requires an edit sequence of $k$ edits to differentiate
%% between this lens and the identity lens.  Simple symmetric lenses avoid this
%% problem by disallowing such lenses: we can't keep track of the number of
%% non-identity edits performed unless there are explicit fields in the data
%% formats keeping track of them.  While we are reducing the expressivity of our
%% lenses with this restriction, we feel the benefits outweigh the costs.
